---
title: "PH 241 Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(epitools)
library(dplyr)
library(knitr)
library(lmtest)
library(epiR)
library(ResourceSelection)
library(foreign)
library(latex2exp) #to add math symbols to the graphs
library(gridExtra) #to arrange plots in a grid
library(esquisse)
```

```{r}
Nhanes_data <- merge(nhanes11_14_2_15_17, nhanes_11_14_sad_vs_wc,by="SEQN") #Merging Datasets

Nhanes_data_select <- select(Nhanes_data, "SEQN", "fincome", "hyper_cat", "hyper_cat2", "race","BMXWT", "BMXHT", "bmipct", "BMI_cat", "RIAGENDR.x" , "age_cat", "PAD680", "INDFMPIR.x", "PA_guide_met.y") #Selecting Variables

Nhanes_data_select <- rename(Nhanes_data_select, RIAGENDR = RIAGENDR.x, INDFMPIR = INDFMPIR.x, PA_guide_met = PA_guide_met.y) #Renaming Variables

Nhanes_data_select
```

```{r}
#manually calculating bmi. not comparable between participants, as ages,sex, and year vary
Nhanes_data_select <- mutate(Nhanes_data_select,BMI = (BMXWT/BMXHT^2)*10000)

#rounding bmi percentiles
Nhanes_data_select <- mutate(Nhanes_data_select, bmipct_rnd = trunc(bmipct, digits = 2))

#creating categoricals of bmi percentiles
Nhanes_data_select <- mutate(Nhanes_data_select, bmipct_cat = ifelse(bmipct_rnd %in% 85:94, 1, ifelse(bmipct_rnd %in% 95:98, 2, ifelse(bmipct_rnd %in% 99:100, 3, 0))))

#creating binary bmi categories
Nhanes_data_select <- mutate(Nhanes_data_select, bmi_bi = ifelse(bmipct_rnd %in% 85:100, 1, 0))

#comment in below if you decide to use these categories for sedentary hours
#Nhanes_data_select <- mutate(Nhanes_data_select, sedentary = ifelse(PAD680 %in% 61:120, 1, ifelse(PAD680 %in% 121:180, 2, ifelse(bmipct_rnd %in% 181:240, 3, ifelse(PAD680 %in% 241:300, 4, ifelse(PAD680 %in% 301:360, 5, ifelse(PAD680 %in% 361:420, 6, ifelse(PAD680 %in% 421:480, 7, ifelse(PAD680 %in% 481:540, 8, ifelse(PAD680 %in% 541:600, 9, ifelse(PAD680 %in% 601:660, 10, ifelse(PAD680 %in% 661:720, 11, ifelse(PAD680 %in% 721:780, 12, ifelse(PAD680 %in% 781:840, 13, ifelse(PAD680 %in% 841:900, 14, ifelse(PAD680 %in% 901:960, 15, 0))))))))))))))))#

#creating binary of sedentary hours
Nhanes_data_select <- mutate(Nhanes_data_select, sedentary = ifelse(PAD680 %in% 481:960, 1, 0))

BMIcat_hyp_4 <- epitable(Nhanes_data_select$bmipct_cat, Nhanes_data_select$hyper_cat2, ncol=2,byrow=T, rev = c("neither")) #2x2 Table Set Up

BMIcat_hyp_4

#rownames(BMIcat_hyp_4) <-c("Under/normal", "Overweight", "Obese I", "Obese II") # Naming Rows
#colnames(BMIcat_hyp_4) <-c("Normal", "Hypertension") # Naming Columns

epitab(BMIcat_hyp_4, method = c("oddsratio")) #Calculating Relative Risk and CIs

```


```{r}
BMIcat_hyp_activity <-with(Nhanes_data_select,table(bmipct_cat, hyper_cat2, by = sedentary))
      
#BMIcat_hyp_RIAGENDR <-with(Nhanes_data_select,table(bmipct_cat, hyper_cat2, by = RIAGENDR))

#BMIcat_hyp_race # 2x2 of BMI vs Hypertension by Race 

#BMIcat_hyp_RIAGENDR # 2x2 of BMI vs Hypertension by Gender 

BMIcat_hyp_activity

sedentary_hyp <- epitable(Nhanes_data_select$sedentary, Nhanes_data_select$hyper_cat2, ncol=2,byrow=T, rev = c("neither"))

sedentary_hyp

mean(Nhanes_data_select$sedentary)
```

```{r}
#Intercept Model (a)

inte_model<-glm(hyper_cat2~1, data = Nhanes_data_select, family = binomial("logit"))

summary(inte_model)$coef

confint(inte_model)

logLik(inte_model)

broom::tidy(inte_model,conf.int = FALSE,exponentiate = TRUE)

```


```{r}
#Intercept + Slope Model

int_slope <- glm(hyper_cat2~bmi_bi, data = Nhanes_data_select, family = binomial("logit"))

summary(int_slope)$coef

confint(int_slope)

lrtest(int_slope)

broom::tidy(int_slope,conf.int = TRUE,exponentiate = TRUE)
```


```{r}
#Indicator Model (bmipct_cat)

Nhanes_data_select <- Nhanes_data_select%>%mutate(Z1 =case_when(bmipct_cat==1~1,
                                                                    TRUE~0),
                                                  Z2 =case_when(bmipct_cat==2~1,
                                                                    TRUE~0),
                                                  Z3 =case_when(bmipct_cat==3~1,
                                                                    TRUE~0))

indi_mod <-glm(hyper_cat2~Z1+Z2+Z3, data = Nhanes_data_select, family = binomial("logit"))

summary(indi_mod)$coef #Model Results + Wald Test

confint(indi_mod) #Confidence Interval

lrtest(int_slope,indi_mod) #Likelihood Ratio Test

broom::tidy(indi_mod,conf.int = TRUE,exponentiate = TRUE)

#THIS TELLS US ITS THRESHOLD

#Goodness of Fit Test
#h1 <- hoslem.test(x= Nhanes_data_select$hyper_cat2,
           # y = predict(int_slope, type = "response", newdata = Nhanes_data_select),
           # g = 2)
#h1

# Test of Association
```

```{r}
# Ordered Model (bmipct_cat); testing for trend in log OR

order_mod <-glm(hyper_cat2~bmipct_cat, data = Nhanes_data_select, family = binomial("logit"))

summary(order_mod)$coef #Model Results + Wald Test

confint(order_mod) #Confidence Interval

lrtest(indi_mod, order_mod) #Likelihood Ratio Test

broom::tidy(order_mod,conf.int = TRUE,exponentiate = TRUE)

```



```{r}
#Confounding Inter-Slope Model

intercept_confound_mod <-glm(hyper_cat2~bmi_bi + sedentary, data = Nhanes_data_select, family = binomial("logit"))

summary(intercept_confound_mod)$coef

confint(intercept_confound_mod)

lrtest(int_slope, intercept_confound_mod)

broom::tidy(intercept_confound_mod,conf.int = TRUE,exponentiate = TRUE)

```

```{r}
#Interaction Model

intercept_interact_mod <-glm(hyper_cat2~bmi_bi*sedentary, data = Nhanes_data_select, family = binomial("logit"))

summary(intercept_interact_mod)$coef

confint(intercept_interact_mod)

lrtest(intercept_confound_mod, intercept_interact_mod)

broom::tidy(intercept_interact_mod,conf.int = TRUE,exponentiate = TRUE)

#h2 <- hoslem.test(x= Nhanes_data_select$hyper_cat2,
          #  y = predict(confound_mod, type = "response", newdata = Nhanes_data_select),
          #  g = 2)
#h2

```


``````{r}
#Evaluate Race as Confounder (Indicator Model)

Nhanes_data_select <- Nhanes_data_select%>%mutate(Z1_race =case_when(race==2~1,
                                                                    TRUE~0),
                                                  Z2_race =case_when(race==3~1,
                                                                    TRUE~0),
                                                  Z3_race =case_when(race==4~1,
                                                                    TRUE~0))

testmodel2 <-glm(hyper_cat2~bmi_bi + Z1_race + Z2_race + Z3_race, data = Nhanes_data_select, family = binomial("logit"))

testmodel3 <-glm(hyper_cat2~bmi_bi + as.factor(race), data = Nhanes_data_select, family = binomial("logit"))

summary(testmodel2)$coef

confint(testmodel2)

lrtest(int_slope,testmodel2)

broom::tidy(testmodel2,conf.int = TRUE,exponentiate = TRUE)

summary(testmodel3)$coef

confint(testmodel3)

lrtest(int_slope,testmodel3)

broom::tidy(testmodel3,conf.int = TRUE,exponentiate = TRUE)
```


```{r}
#Evaluate Race as Interaction Term

testmodel4 <-glm(hyper_cat2~bmi_bi*as.factor(race), data = Nhanes_data_select, family = binomial("logit"))

summary(testmodel4)$coef

confint(testmodel4)

lrtest(testmodel3,testmodel4)

broom::tidy(testmodel4,conf.int = TRUE,exponentiate = TRUE)

```


```{r}
#Indicator with Confounding Model

ind_confound_mod <-glm(hyper_cat2~as.factor(bmipct_cat)+sedentary, data = Nhanes_data_select, family = binomial("logit"))

summary(ind_confound_mod)$coef

confint(ind_confound_mod)

lrtest(indi_mod, ind_confound_mod)

h2 <- hoslem.test(x= Nhanes_data_select$hyper_cat2,
            y = predict(ind_confound_mod, type = "response", newdata = Nhanes_data_select),
            g = 2)
h2
```


```{r}
#GRAPHS OF RISK, ODDS, AND LOG ODDS FOR bmipct_cat Ordered Model

#Estimate the risk using the summarize function
risk_table <- Nhanes_data_select %>% 
  group_by(bmipct_cat) %>% 
  summarize(risk = mean(hyper_cat2))

risk_table <- risk_table %>%
  mutate(odds = risk/(1-risk)) %>% #estimate the odds 
  mutate(log.odds = log(odds)) #estimate the log odds

#Generate the model-based risk estimates
Nhanes_data_select$pred_risk <- predict(int_slope, #use the model to...
                                   type = "response", #predict the risk for...
                                   newdata = Nhanes_data_select) #the Nhanes data

#Generate the model-based odds and log odds estimates
Nhanes_data_select <- Nhanes_data_select %>%
  mutate(pred_odds = pred_risk/(1-pred_risk)) %>%
  mutate(pred_log_odds = log(pred_odds))

risk_plot <- risk_table %>%
  #the observed data from before
  ggplot(aes(x = bmipct_cat, y = risk)) + 
  geom_point(alpha = 0.5) + 
  #the fitted line
  geom_line(aes(y = pred_risk),
            data = Nhanes_data_select,
            col = "blue") + 
  #general formatting
  scale_x_continuous("Body Mass Index", breaks = 0:3, labels =  
                            c("Under/Normal",
                              "Overweight",
                              "Obese I",
                              "Obese II")) +  
  ylab("Risk of Hypertension") + 
  theme_classic()

odds_plot <- risk_table %>%
  #the observed data from before
  ggplot(aes(x = bmipct_cat, y = odds)) + 
  geom_point(alpha = 0.5) + 
  #the fitted line
  geom_line(aes(y = pred_odds),
            data = Nhanes_data_select,
            col = "blue") + 
  #general formatting
  scale_x_continuous("Body Mass Index", breaks = 0:3, labels = 
                            c("Under/Normal",
                              "Overweight",
                              "Obese I",
                              "Obese II")) +  
  ylab("Odds of Hypertension") + 
  theme_classic()

log_odds_plot <- risk_table %>%
  #the observed data from before
  ggplot(aes(x = bmipct_cat, y = log.odds)) + 
  geom_point(alpha = 0.5) + 
  #the fitted line
  geom_line(aes(y = pred_log_odds),
            data = Nhanes_data_select,
            col = "blue") + 
  #general formatting
  scale_x_continuous("Body Mass Index", breaks = 0:3, labels = 
                            c("Under/Normal",
                              "Overweight",
                              "Obese I",
                              "Obese II")) +
                                
  ylab("Log odds of Hypertension") + 
  theme_classic()

#grid.arrange(
  #risk_plot,
  #odds_plot,
  #log_odds_plot,
  #ncol = 2)

log_odds_plot
```


```{r}
#Plotting BMI log OR graphs by Sedentary Activity (as Binary)
preds_con <- data.frame(id_con = Nhanes_data_select$SEQN, 
                    bmi_con = Nhanes_data_select$bmipct_cat,
                    sedentary_con = as.factor(Nhanes_data_select$sedentary),
                    pred_risk_con =  predict(intercept_confound_mod,
                 newdata = Nhanes_data_select,
                 type = "response"))


preds_con <- preds_con %>%
  mutate(odds = pred_risk_con/(1-pred_risk_con)) %>%
  mutate(log_odds = log(odds))

p1 <- preds_con %>%
  mutate(sedentary_con = ifelse(sedentary_con == 1, "More than Six Hours", "Less than Six Hours")) %>%
  ggplot(aes(x = bmi_con, y = log_odds, col = sedentary_con)) + 
  geom_line() + 
  scale_x_continuous("Body Mass Index", breaks = 0:3, labels = 
                            c("Under/Normal",
                              "Overweight",
                              "Obese I",
                              "Obese II")) +  
  theme(legend.title = element_blank()) +  #hide legend title
  ylab("Log odds of Hypertension") +
  ggtitle("Model Results: Hypertension ~ BMI + Hours of Sedentary Activity", subtitle = "This assumes no interaction, which is why the lines are parallel.")

p1
```

# GRAPHS OF RISK, ODDS, AND LOG ODDS FOR BMIPCT_CAT ORDERED MODEL

```{r}
#Estimate the risk using the summarize function
risk_table_5 <- Nhanes_data_select %>% 
  group_by(bmipct_cat) %>% 
  summarize(risk = mean(hyper_cat2))

risk_table_5 <- risk_table_5 %>%
  mutate(odds = risk/(1-risk)) %>% #estimate the odds 
  mutate(log.odds = log(odds)) #estimate the log odds

#Generate the model-based risk estimates
Nhanes_data_select$pred_risk <- predict(mod3_bmipct_cat, #use the model to...
                                   type = "response", #predict the risk for...
                                   newdata = Nhanes_data_select) #the Nhanes data

#Generate the model-based odds and log odds estimates
Nhanes_data_select <- Nhanes_data_select %>%
  mutate(pred_odds = pred_risk/(1-pred_risk)) %>%
  mutate(pred_log_odds = log(pred_odds))

risk_plot_5 <- risk_table_5 %>%
  #the observed data from before
  ggplot(aes(x = bmipct_cat, y = risk)) + 
  geom_point(alpha = 0.5) + 
  #the fitted line
  geom_line(aes(y = pred_risk),
            data = Nhanes_data_select,
            col = "blue") + 
  #general formatting
  scale_x_continuous("Body Mass Index", breaks = 0:3, labels = 
                            c("Under/Normal",
                              "Overweight",
                              "Obese I",
                              "Obese II")) +  
  ylab("Risk of Hypertension") + 
  theme_classic()

odds_plot_5 <- risk_table_5 %>%
  #the observed data from before
  ggplot(aes(x = bmipct_cat, y = odds)) + 
  geom_point(alpha = 0.5) + 
  #the fitted line
  geom_line(aes(y = pred_odds),
            data = Nhanes_data_select,
            col = "blue") + 
  #general formatting
  scale_x_continuous("Body Mass Index", breaks = 0:3, labels = 
                            c("Under/Normal",
                              "Overweight",
                              "Obese I",
                              "Obese II")) +  
  ylab("Odds of Hypertension") + 
  theme_classic()

log_odds_plot_5 <- risk_table_5 %>%
  #the observed data from before
  ggplot(aes(x = bmipct_cat, y = log.odds)) + 
  geom_point(alpha = 0.5) + 
  #the fitted line
  geom_line(aes(y = pred_log_odds),
            data = Nhanes_data_select,
            col = "blue") + 
  #general formatting
  scale_x_continuous("Body Mass Index", breaks = 0:3, labels = 
                            c("Under/Normal",
                              "Overweight",
                              "Obese I",
                              "Obese II")) +
                                
  ylab("Log odds of Hypertension") + 
  theme_classic()

grid.arrange(
  risk_plot_5,
  odds_plot_5,
  log_odds_plot_5,
  ncol = 2)
```

# Other models evaluated:

```{r}
#Ordered Model (BMI) (continuous)

mod3_BMI <-glm(hyper_cat2~BMI, data = Nhanes_data_select, family = binomial("logit"))

summary(mod3_BMI)$coef #Model Results + Wald Test

lrtest(mod3_BMI) #Likelihood Ratio Test

confint(mod3_BMI) #Confidence Interval
```


```{r}
#Ordered Model (bmipct) (continuous)

mod3_bmipct <-glm(hyper_cat2~bmipct, data = Nhanes_data_select, family = binomial("logit"))

summary(mod3_bmipct)$coef #Model Results + Wald Test

lrtest(mod3_BMI, mod3_bmipct) #Likelihood Ratio Test

confint(mod3_bmipct) #Confidence Interval

#esquisse::esquisser(Nhanes_data_select)
```






